Boutiques: a flexible framework for automated application integration in computing platforms

### 2017-10-30 boutiquespaper.tex
Worked on introduction

\abstract{Porting applications to execution platforms such as web
  portals and workflow engines is critical to enable their sharing
  among scientific communities and their efficient execution on
  computing infrastructures. However, application porting remains a
  costly human effort that consists of 1)~installing the application
  on the target infrastructure, 2)~describing the application in a
  format compatible with the execution platform, and 3)~generating
  proper user interfaces. Due to the variety of computing platforms,
  application porting efforts are often replicated several times while
  interoperability would save cost and improve the quality of the ported
  applications. We describe \boutiques, a descriptive framework and
  command line utility that allows for automatic import and exchange of
  applications across platforms. \boutiques adopts a rich, flexible
  application description format which relies on Linux container
  environments such as Docker and Singularity to install applications
  across platforms in a lightweight manner. \boutiques is currently used
  by three distinct virtual research platforms, and has been used to
  describe dozens of neuroinformatics tools.}

@@

Platforms such as web services, portals, and virtual research
environments are becoming an common solution for increasing the
accessibility of high performance computing. These scientific gateways and
workflow engines often rely on third-party data processing and simulation
applications to deliver services to their users. Such applications are
often repeatedly and manually ported to various platforms whereas this
porting effort would greatly benefit from being mutualized. Meanwhile,
virtual container systems such as Docker and Singularity greatly facilitate
the sharing of software and improve the reproducibility of analyses by defining
immutable, reusable execution environments. However, containers need
to be properly annotated to allow platforms to leverage the software
they contain. The command lines that can be run in a container must be formally
described so that they can be automatically created and executed by platforms.
Describing command lines is not straightforward since they often contain
implicit semantics specified in human-readable documents such as manuals.

\boutiques is a system to share command-line applications across
platforms. In \boutiques, a command line is described using a flexible
template comprising the inputs it requires and the outputs it
produces. Inputs may be inter-dependent, e.g., input \emph{A} cannot be used
if input \emph{B} is used; inputs may be passed in configuration files
or directly on the command line. Such formal descriptions, simply referred to
as \emph{descriptors}, allow for advanced validation of input values to
prevent errors. A descriptor refers to a container where a given application
and all their dependencies are installed. \boutiques also provides a
simple parallelization framework that allows applications such as
workflow engines to concurrently execute multiple tasks (in the
remainder, tasks are called \emph{invocations}). A set of core tools
facilitate the construction, validation, import, execution, and publishing
of \boutiques descriptors. \boutiques descriptors are intented to be produced by
scientific developers and stored alongside their application, indexed by common
repositories, and to be consumed by execution platforms.

>>>>

\abstract{Porting applications to execution platforms such as web
  portals and workflow engines is critical to enable their sharing
  among scientific communities and their efficient execution on
  computing infrastructures. However, application porting remains a
  costly human effort that consists of 1)~installing the application
  on the target infrastructure, 2)~describing the application in a
  format compatible with the execution platform, and 3)~generating
  proper user interfaces. Due to the variety of computing platforms,
  application porting efforts are often replicated several times while
  interoperability would save cost and improve the quality of the ported
  applications. We describe \boutiques, a descriptive framework and
  command line utility that allows for automatic import and exchange of
  applications across platforms. \boutiques adopts a rich, flexible
  application description format which relies on Linux container
  environments Docker and Singularity to install applications
  across platforms in a lightweight manner. \boutiques is currently used
  by three distinct virtual research platforms, and has been used to
  describe dozens of neuroinformatics tools.}

@@

Platforms such as web services, portals, scientific gateways, workflow
engines and virtual research environments commonly integrate
third-party applications to facilitate the processing of large
datasets on high-performance computing resources. Such applications,
however, are often repeatedly and manually ported to various platforms
whereas this porting effort would greatly benefit from being
mutualized. Meanwhile, virtual container systems such as Docker and
Singularity have emerged to facilitate the sharing of software and
improve the reproducibility of analyses by defining immutable,
reusable execution environments. Nevertheless, the automatic
integration of containerized applications in platforms require formal
annotations of container images, to describe the available command lines.

\boutiques is a system to share command-line applications across
platforms. In \boutiques, a command line is described using a flexible
template comprising the inputs it requires and the outputs it
produces. Inputs may be inter-dependent and passed in configuration
files or directly on the command line. Such formal descriptions,
simply referred to as \emph{descriptors}, link to a container image
where the given application and all its dependencies are
installed. \boutiques descriptors allow for automatic application
integration in platforms, and advanced validation of input values to
prevent errors. \boutiques also provides a simple parallelization
framework that allows parallel applications to concurrently execute
multiple tasks (in the remainder, tasks are called
\emph{invocations}). \boutiques descriptors are intented to be
produced by scientific application developers, stored alongside their
application, indexed by common repositories, and consumed by
execution platforms.  A set of core tools facilitate the construction,
validation, import, execution, and publishing of \boutiques
descriptors.
